Volunteers in the Clouds: an Architecture for Low Cost and Potentially Massive Distributed Evolutionary Computation

### 2015-07-31 volunteer-clouds-ieee-ci-si.tex
Provisional results

In general, a volunteer-based distributed evolutionary computation
system based on the browser is simply a client-server architecture
whose client is, or can be, embedded in the browser via
JavaScript. All parts of the system are free and available with a free
license from \url{https://github.com/JJ/splash-volunteer}.
Thus, {\sf NodIO} has two parts:\begin{enumerate}
\item A REST server, that is, a server that includes several remote
  requests for storing and retrieving information (the CRUD cycle:
  create, request, update and delete) from the server. The information
  is of two kinds: {\em problem} based, that is, related to the
  evolutionary algorithm such as {\tt PUT}ing a chromosome in or {\tt
    GET}ing a random chromosome from it. It also peforms logging
  duties, but is basically a very lightweight and hign performance
  data store \cite{jj:idc:lowcost}. The server has the capability to
  run a single experiment, storing the chromosomes in a data structure
  that is reset when the solution is found. 
\item A client that includes the evolutionary algorithm as 
  JavaScript code embedded in a web page that displays graphs and some
  additional links and information on the experiment. This code runs
  an evolutionary algorithm {\em island} starting with random
  population and sending, every 100 generations, the best individual
  back to the server (via a {\tt PUT} request) and requesting a random
  individual from the server (via a {\tt GET} request). Apart from
  these constants, the algorithm can be run in many different ways:
  stopping when solution is found or, as we have done in this paper,
  using WebWorkers, which run in the background even if the webpage
  with the experiment is not currently on focus. 
\end{enumerate}

@@

>>>>

In general, a volunteer-based distributed evolutionary computation
system based on the browser is simply a client-server architecture
whose client is, or can be, embedded in the browser via
JavaScript. All parts of the system are free and available with a free
license from \url{https://github.com/JJ/splash-volunteer}.
Thus, {\sf NodIO} has two parts:\begin{enumerate}
\item A REST server, that is, a server that includes several remote
  requests for storing and retrieving information (the CRUD cycle:
  create, request, update and delete) from the server. The information
  is of two kinds: {\em problem} based, that is, related to the
  evolutionary algorithm such as {\tt PUT}ing a chromosome in or {\tt
    GET}ing a random chromosome from it. It also peforms logging
  duties, but is basically a very lightweight and hign performance
  data store \cite{jj:idc:lowcost}. The server has the capability to
  run a single experiment, storing the chromosomes in a data structure
  that is reset when the solution is found. 
\item A client that includes the evolutionary algorithm as 
  JavaScript code embedded in a web page that displays graphs and some
  additional links and information on the experiment. This code runs
  an evolutionary algorithm {\em island} starting with random
  population and sending, every 100 generations, the best individual
  back to the server (via a {\tt PUT} request) and requesting a random
  individual from the server (via a {\tt GET} request). Apart from
  these constants, the algorithm can be run in many different ways:
  stopping when solution is found or, as we have done in this paper,
  using WebWorkers, which run in the background even if the webpage or
  browser tab with the experiment is not currently on focus.  
\end{enumerate}

@@

First we establish the baseline performance by running the
evolutionary algorithm in a desktop client written using {\sf NodEO},
the node.js evolutionary algorithm library. This experiment tries to
find the solution to the 40-trap function with parameters $l=4, a=1,
b=2, z=3$, a population size equal 512, which was found in previous
experiments to be the correct one for finding the solution 95\% of the
time. The algorithm was run until the solution (a string with all
ones) was found or five millions evaluations had been performed. It
took around a minute, on average, that is $68.9694$ seconds, to
perform the fifty runs.
